---
title: "Programming Assignment 4"
subtitle: "Data Science for Linguists"
author: "Meritxell Feliu Ribas"
institute: "Rutgers University"
date: "2023/03/27 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: [default, rutgers, rutgers-fonts]
---

# Load Packages

```{r, load-packages}
library(tidyverse)
library(ds4ling)
library(here)
library(broom)
```

---

# Load Data

```{r, load-data}
fricatives_data <- read.csv(here("data_raw","fricatives","fricatives.csv"))
```

---

# Tidy Data

```{r, tidy-data} 
fricatives_tidy <- fricatives_data %>%
  pivot_longer(
    cols = c("s_cog", "sh_cog", "s_skewness", "sh_skewness"),
    names_to = "cog_skewness",
    values_to = "values"
  ) %>%
  separate(
    col = "cog_skewness", 
           into = c("phoneme", "cog_skewness"), 
           sep = "_"
  ) %>%
  pivot_wider(
    names_from = "cog_skewness",
    values_from = "values"
  ) %>%
  write_csv(here("data_tidy", "fricatives_tidy.csv"))
```

---

# Descriptive Statistics

```{r, descriptive-statistics}
fricatives_tidy %>%
  group_by(phoneme) %>% 
  summarize(
    mean_cog = mean(cog),
    sd_cog = sd(cog),
    mean_skewness = mean(skewness),
    sd_skewness = sd(skewness),
    ) %>%
  knitr::kable()
```

---

# Plots

### Center of gravity as a function of phoneme

```{r, cog-as-a-function-of-phoneme, fig.height=4}
fricatives_tidy %>% 
 ggplot() + 
  aes(x = phoneme, y = cog) + 
  geom_boxplot() +
  labs(x = "Phoneme", y = "Center of Gravity",
       title = "Center of Gravity by Phoneme")
```

---

### Skewness as a function of phoneme with summary

```{r, skewness-as-a-function-of-phoneme-with-summary, fig.height=5}
fricatives_tidy %>%
  ggplot() +
  aes(x = phoneme, y = skewness) +
  geom_point() +
  stat_summary(fun.data = "mean_se", geom = "point") +
  labs(x = "Phoneme", y = "Skewness",
       title = "Skewness by Phoneme")
```

---

# Model
```{r, cog-as-a-function-of-skewness-s}
# Data transformation
phoneme_s <- filter(fricatives_tidy, phoneme == "s")

# Model
model_0 <- lm(cog ~ 1, data = phoneme_s)
model_s <- lm(cog ~ skewness, data = phoneme_s)

#Nested model comparisons
anova(model_0, model_s)
```

---

```{r}
summary(model_s)

tidy(model_s) %>%
  knitr::kable(format = "pandoc")
```

---

# Scatter plot of model_s
```{r, scatterplot-model_s, fig.height=4}
phoneme_s %>% 
  ggplot() + 
  aes(x = skewness, y = cog) + 
  geom_point() +
  geom_smooth(method = lm) +
  labs(y = "Center of Gravity", x = "Skewness",
       title = "Center of Gravity by Skewness of the [s] Phoneme")
```

---

# Model Diagnosis

```{r, model-diagnosis, fig.height=6}
diagnosis(model_s)
```

---

# Results
The data were analyzed using a linear regression model to examine the relationship between the center of gravity and the skewness of the phoneme [s]. The model included *center of gravity* as a dependent variable and *skewness* as a predictor. Main effects were assessed using nested model comparisons. Experiment-wise alpha was set at 0.05. 

A linear relationship between skewness and center of gravity can be observed. In our model, nearly 88% of the variance found in center of gravity can be explained by skewness (adjusted R-squared = 0.8771). Specifically, a one unit increase in skewness is associated with an increase of 3029.1 in center of gravity (SE = 259.1, t = 11.69, p < 0.001). The data is approximately normal distributed based on the Q-Q plot. There is a main effect of skewness (F(1) = 136.64, p < 0.001). 

---

# Assumptions

```{r, load-data-assumptions}
assumptions_data <- read.csv(here("data_raw","fricatives","assumptions.csv"))
```

```{r, scatterplot-assumptions, fig.width=6, fig.height=4}
assumptions_data %>% 
  ggplot() + 
  aes(x = x, y = y) + 
  geom_point() +
  labs(title = "Assumptions")
```

(Thanks Joseph, this dinosaur made my day!)

---
### Comment about the assumptions dataset

The assumptions dataset would not be appropriate to fit this to a linear model because the relationship between the data is not linear, as we can see in the scatter plot (it is a dinosaur!). There may be a relationship between these two variables, but a linear model is not a suitable method to analyze them. 

